The problem and promise of scale in animal social multilayer networks

### 2020-02-09 paper/manuscript.Rmd
more refz

In contrast to spatial networks, where nodes represent spatial locations that
are fixed in space, movement networks can represent continuous and simultaneous
movement trajectories of multiple animals. In a recent study of vulturine 
gunieafowl (Acryllium vulturinum), GPS movement data were used to generate 
monolayer association networks within a multilayer network [@Papageorgiou_2019]. 
However, despite potential to generate networks in different spatial or social contexts, 
layers in the multilayer network represented social associations within and 
between groups. Continuous collection of animal relocation data using GPS 
technology can be used to measure how groups of animals make collective 
movement decisions in the context of their local environment (Strandburg-Peshkin et al. 2015, 2017) and across scales. A
collective movement layer could constitute similarity in movement for dyads
[@Long_2014] or groups  [@Bode_2011]. This movement layer could be
made at a fine (i.e. minute-to-minute: [@Cleasby_2019] or coarse scale
(global migration: Flack et al. 2018). Our case study integrates variation 
in temporal and spatial scale to assess the role of scale in predicting 
multilayer networks. As one might expect, we found increasing spatial 
scale results in saturation of potential social connections (Figure 3), 
while increasing temporal scale had little effect on habitat-specific 
networks (Figure 5). Another potential application would be to
generate social networks during times when animals are engaged in different
behaviours, including traveling, interacting, or foraging (e.g. Muller et al.
2018; Finn et al. 2019) and construct a multilayer network based on
behaviour-specific monolayer networks (e.g. Smith-Aguilar et al. 2019).

Importantly, the key to incorporating movement in multilayer networks across
scales is the technological advancements available to overcome previous
limitations (Hughey et al. 2018). Biologging and wildlife tracking technology 
has advanced to the point where relocations can be programmed to occur as 
frequently as each second (Börger 2016), a practice which has been applied 
to questions of collective movement (e.g Strandburg-Peskin et al. 2015). 
Meanwhile, remote sensing technology, including unmanned aerial vehicles 
and satellites, can be used to track the fine-scale movements of groups 
over large distances and time scales (Torney et al. 2019). Using 
technology to infer behaviour has therefore become standard operating procedure
for many species in many systems. Our ability to collect and process 
remotely sensed data and infer social processes based on animal movement 
provides an opportunity to use multilayer social network analysis to answer
complex questions about animal socioecology.

Given the existing framework for constructing multilayer networks from
behavioural layers (e.g. Smith-Aguilar et al. 2019), monolayer networks that
explicity quantify movement paramters could be itnegrated. For example, edges 
could represent measures of dyadic or collective movement based on dynamic 
interaction metrics [@Long_2014]. Integrating movement layers
with habitat-specific social association layers could further shed light on the
role of movement as a driver of the social and spatial environments an
individual experiences [@Webber_2018]. In our case study, we
generated multilayer networks based on habitat-specific monolayers. Coordinated
movement of dyads, or groups, could vary based on habitat, and the approach we
outline could be used to explore mechanisms linking the social and spatial
environments. For example, social processes may be an emergent property of the
landscape because animals aggregate at resources, such as waterholes
(Chamaillé-Jammes et al. 2008). Movement to, from, or within habitat patches  or
territories can also contribute to the formation of the social environment
(Spiegel et al. 2016). For example, in sleepy lizards (*Tiliqua rugosa*) males
interacted primarily at home range edges, presumably driving territorial
behaviour (Spiegel et al. 2018). Incorporating movement ecology within a
multilayer network framework is a logical next step for a burgeoning field. In
particular, when association networks are habitat-specific, we suggest the
integration of movement and space use in multilayer networks provides novel
insights into the effects of habitat configuration on the formation of the
social environment [@He_2019].

>>>>

In contrast to spatial networks, where nodes represent spatial locations that
are fixed in space, movement networks can represent continuous and simultaneous
movement trajectories of multiple animals. In a recent study of vulturine 
gunieafowl (Acryllium vulturinum), GPS movement data were used to generate 
monolayer association networks within a multilayer network [@Papageorgiou_2019]. 
However, despite potential to generate networks in different spatial or social contexts, 
layers in the multilayer network represented social associations within and 
between groups. Continuous collection of animal relocation data using GPS 
technology can be used to measure how groups of animals make collective 
movement decisions in the context of their local environment (Strandburg-Peshkin et al. 2015, 2017) and across scales. A
collective movement layer could constitute similarity in movement for dyads
[@Long_2014] or groups  [@Bode_2011]. This movement layer could be
made at a fine [i.e. minute-to-minute: @Cleasby_2019] or coarse scale
[global migration: @Flack_2018]. Our case study integrates variation 
in temporal and spatial scale to assess the role of scale in predicting 
multilayer networks. As one might expect, we found increasing spatial 
scale results in saturation of potential social connections (Figure 3), 
while increasing temporal scale had little effect on habitat-specific 
networks (Figure 5). Another potential application would be to
generate social networks during times when animals are engaged in different
behaviours, including traveling, interacting, or foraging [@Muller_2018; @Finn_2019] 
and construct a multilayer network based on behaviour-specific monolayer 
networks [e.g. @Smith_Aguilar_2019].

Importantly, the key to incorporating movement in multilayer networks across
scales is the technological advancements available to overcome previous
limitations [@Hughey_2018]. Biologging and wildlife tracking technology 
has advanced to the point where relocations can be programmed to occur as 
frequently as each second [@Börger_2016], a practice which has been applied 
to questions of collective movement (e.g Strandburg-Peskin et al. 2015). 
Meanwhile, remote sensing technology, including unmanned aerial vehicles 
and satellites, can be used to track the fine-scale movements of groups 
over large distances and time scales (Torney et al. 2019). Using 
technology to infer behaviour has therefore become standard operating procedure
for many species in many systems. Our ability to collect and process 
remotely sensed data and infer social processes based on animal movement 
provides an opportunity to use multilayer social network analysis to answer
complex questions about animal socioecology.

Given the existing framework for constructing multilayer networks from
behavioural layers [e.g. @Smith_Aguilar_2019], monolayer networks that
explicity quantify movement paramters could be itnegrated. For example, edges 
could represent measures of dyadic or collective movement based on dynamic 
interaction metrics [@Long_2014]. Integrating movement layers
with habitat-specific social association layers could further shed light on the
role of movement as a driver of the social and spatial environments an
individual experiences [@Webber_2018]. In our case study, we
generated multilayer networks based on habitat-specific monolayers. Coordinated
movement of dyads, or groups, could vary based on habitat, and the approach we
outline could be used to explore mechanisms linking the social and spatial
environments. For example, social processes may be an emergent property of the
landscape because animals aggregate at resources, such as waterholes
(Chamaillé-Jammes et al. 2008). Movement to, from, or within habitat patches  or
territories can also contribute to the formation of the social environment
(Spiegel et al. 2016). For example, in sleepy lizards (*Tiliqua rugosa*) males
interacted primarily at home range edges, presumably driving territorial
behaviour (Spiegel et al. 2018). Incorporating movement ecology within a
multilayer network framework is a logical next step for a burgeoning field. In
particular, when association networks are habitat-specific, we suggest the
integration of movement and space use in multilayer networks provides novel
insights into the effects of habitat configuration on the formation of the
social environment [@He_2019].
